Embedding Dimension of Polyhedral Losses

### 2020-01-31 colt-20/embed-dim.tex
Cleaning appendix of polytope notes and lemma proofs before class

\newcommand{\Opt}{\mathrm{Opt}}
\newcommand{\risk}[1]{\underline{#1}}
\newcommand{\inprod}[2]{\langle #1, #2 \rangle}%\mathrm{int}(#1)}
\newcommand{\inter}[1]{\mathrm{int}(#1)}%\mathrm{int}(#1)}
%\newcommand{\expectedv}[3]{\overline{#1}(#2,#3)}
\newcommand{\expectedv}[3]{\E_{Y\sim{#3}} {#1}(#2,Y)}
\newcommand{\toto}{\rightrightarrows}
\newcommand{\strip}{\mathrm{strip}}
\newcommand{\trim}{\mathrm{trim}}
\newcommand{\fplc}{finite-piecewise-linear and convex\xspace} %xspace for use in text
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\indopp}{\bar{\mathbbm{1}}}
\newcommand{\ones}{\mathbbm{1}}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}

@@

%\begin{theorem}[\cite[Theorem 3.1.2]{weibel2007minkowski}] \label{thm:unique-face-decomp}
%	Let $T_1, \ldots, T_n$ be polytopes in $\reals^d$ and let $F$ be a face of the Minkowski sum $T := T_1 \oplus \ldots \oplus T_n$.
%	Then there are faces $F_1, \ldots, F_n$ of $T_1, \ldots, T_n$ respectively such that $F = F_1 \oplus \ldots \oplus F_n$.
%	Moreover, this decomposition is unique.
%\end{theorem}

\iffalse 
\begin{corollary}[\cite[Corollary 3.1.3]{weibel2007minkowski}]\label{cor:face-decomp-normal-cones}
	Let $T = T_1 \oplus \ldots \oplus T_n$ be a Minkowski sum of polytopes in $\reals^d$, let $F$ be a nonempty face of $T$ , and let $F_1, \ldots, F_n$ be its decomposition.
	Then $\N(F;T) = \N(F_1;T_1) \cap \ldots \cap \N(F_n; T_n)$.
\end{corollary}

%used in proving completeness of constructed V
\begin{corollary}[\cite[Corollary 3.1.4]{weibel2007minkowski}]
	Let $F_1, \ldots, F_n$ be nonempty faces of the polytopes $T_1, \ldots, T_n$ respectively, then $F_1 \oplus \ldots \oplus F_n$ is a face of $T_1 \oplus \ldots \oplus T_n$ if and only if the intersection of their normal cones is nonempty.
\end{corollary}
\fi

%\begin{theorem}[\cite[Theorem 3.1.6]{weibel2007minkowski}]\label{thm:support-minksum}
%	The supporting function of a Minkowski sum is the sum of the supporting functions of its summands.
%\end{theorem}

\begin{definition}[Weighted Minkowski sum]
	If $T_1, \ldots, T_n$ are polytopes in $\reals^d$, we can call $T(\vec p)$ their \emph{weighted} Minkowski sum for $\vec p \in \reals^n_{\geq 0}$,
	\begin{align*}
	T(\vec p) &:= \oplus_y p_y T_y = p_1 T_1 \oplus \ldots \oplus p_n T_n
	\end{align*}
\end{definition}

%From the thesis:  \emph{``It is easy to see that the normal fan (undefined here, but consequently normal cones) of $p_i T_i$ does not change as long as $p_i$ is positive.  Since the normal fan of a Minkowski sum can be deduced from that of its summands, we can deduce from this that the conbimatorial properties of $\oplus_y p_y T_y$ stay the same as long as all $p_i$ are positive.''}

Suppose we are given a polytope $T_y \in \reals^d$ and set of vectors $V \in \reals^{k \times d}$.
Call $e^y \in \reals^k$ the vector such that $e^y_i = \max_{x \in T_y}\inprod{v_i}{x}$.  
For a finite set $\T = \{T_1, , \ldots, T_n\}$, let us denote the \emph{support matrix} $E = (e^y)_{y=1}^n$.
\begin{definition}
	We say a set of normals $V$ is \emph{complete} with respect to a polytope $T_y$ if $T_y = \{x \in \reals^d: Vx \leq e^y\}$.
\end{definition}
Moreover, we say $V$ is complete with respect to the set of polytopes $\T$ if and only if $V$ is complete with respect to each $T_y \in \T$.

\subsubsection*{Optimality Condition}
In general, we know that $\vec 0$ in the subgradient of $f(r)$ if and only if $r$ is an optimum of convex $f$ in general dimensions.

For nice\footnote{for reports on the relative interior of the domain, and something else} conditions, we have $\vec 0 \in \oplus_i f_i$ if and only if $\vec 0 \in \oplus_i \partial f_i(r)$.
This allows us to consider the following optimality condition:
\begin{definition}\label{def:optimality}[Optimality]
	For all $r \in \R$ and $y \in \Y$, we have $r \in \gamma_r \iff \vec 0 \in \oplus_y p_y T_y(r)$.
\end{definition}
The polytope $T_y(r)$ can be thought of as $\partial L_y(\varphi(r))$.

@@

\subsection{Given normals $V$}
Now, for a given polytope $T(p)$, we want to ask when a given $z \in \reals^d$ is in the polytope $T(p)$.
We will later generalize to finding the set of $p \in \simplex$ for which $\vec 0 \in T(p)$ by substituting $z= \vec 0$.
Throughout, assume we have $V$ which is complete for $\T$ and $E$ defined by the support of each normal in $V$ for all $T_y \in \T$.
We denote $e^y = E_{;y}$ as the $y^{th}$ column of $E$, or equivalently, the support vector for $T_y$ given $V$.

@@

\subsection{Equivalences}
In the big picture, we want to know two things.
First, if we are given a cell $C$, how we can construct a finite set of polytopes $\T = \{T_1, \ldots, T_n\}$ such that $\vec 0 \in T(p) \iff p \in C$.
Second, we want to know the opposite case: starting from $\T$, can we derive the cell $C \subseteq \simplex$ where $\vec 0 \in T(p)$ for all $p \in C$?
We start with the latter, leaving the former for future work.

We know that if we are given $\T$ and a complete set of normals $V$, we can describe $D = \{p \in \simplex : Ep \geq \vec 0\}$ as in Section~\ref{sec:start-polytope}.
However, we do not necessarily have $D' := \{p \in A : Ep \geq \vec 0\} \subseteq \simplex$, and this poses some issues given the construction of the cell $C$ only assuming that we are in the affine span of the simplex, but not necessarily in the simplex.

@@

\section{$d$-dimensional omitted proofs}\label{app:omitted-d-dim}
\begin{proof}[Proof of Lemma~\ref{lemma:minkowski-support}]
	\jessie{Appendix}
	For each $p$, $\oplus_p \T$ is a polytope.
	\bo{TODO: it turns out that $p$ doesn't matter, only its support; and there are only finitely many possible supports (or even better, the normals that are complete for the uniform minkowski sum turn out to be complete for all other $p$-combinations).}
	For each of the finitely many supports ($2^n - 1$), we know $\oplus_p \T$ is a polytope, and every polytope can be uniquely defined by an enumeration of vectors normal to its facets.
	As a two polytopes with the same support can be defined by the same facet enumeration, we can simply concatenate these finite set of normals for the finite polytope supports.
	Vectors that are not normal to a facet simply become redundant constraints.
	This yields finitely many normals defining the weighted Minkowski sum $\oplus_p \T$ for all $p \in \simplex$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lemma:D-polytope}]
	\jessie{Appendix}
	Recall by definition, the notation $\oplus_p \T = \{\sum_y p_y x_y : x_y \in T_y (\forall y)\}$.
	Each $T_y$ is a polytope, so $p_y T_y$ is a polytope.
	The Minkowski sum of polytopes is a polytope, so $\oplus_p \T$ is a polytope by~\cite[Section 1.2]{weibel2007minkowski}.
	\bo{TODO}\jessie{Taking a try}
	Since $\oplus_p \T$ is a polytope for all $p \in \simplex$, we know there is a halfspace representation of normals $V$ so that for all $y \in \Y$, we have $x \in p_yT_y \iff \inprod{V}{x} \leq p_y e^y$ for some matrix $V$ and the support vector $e^y$, where $e^y_i = H_{T_y}(V_i)$.
	Adding redundant constraints for the $e^y$ as necessary, we know that the support of the Minkowski sum for a given normal is the sum of the normals~\cite[Theorem 3.1.6]{weibel2007minkowski}.
	Concatenating the vectors $e^y$ into $E$, we have $x \in \oplus_p \T \iff \inprod{V^*}{x} \leq E^*p$, where $V^*$ has redundant rows added as necessary. \jessie{Hand wavy... not a fan.}
	Substituting $x = \vec 0$, we see $\vec 0 \in \oplus_p \T \iff Ep \geq \vec 0$, which defines a polytope by construction of $E$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lemma:E-to-B}]
	\bo{? Cite} \jessie{Is left as an exercise in the following text:} \cite[Exercise 2.15]{ziegler2012lectures}
\end{proof}

>>>>

\newcommand{\Opt}{\mathrm{Opt}}
\newcommand{\risk}[1]{\underline{#1}}
\newcommand{\inprod}[2]{\langle #1, #2 \rangle}%\mathrm{int}(#1)}
\newcommand{\inter}[1]{\mathrm{int}(#1)}%\mathrm{int}(#1)}
%\newcommand{\expectedv}[3]{\overline{#1}(#2,#3)}
\newcommand{\expectedv}[3]{\E_{Y\sim{#3}} {#1}(#2,Y)}
\newcommand{\toto}{\rightrightarrows}
\newcommand{\strip}{\mathrm{strip}}
\newcommand{\trim}{\mathrm{trim}}
\newcommand{\fplc}{finite-piecewise-linear and convex\xspace} %xspace for use in text
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\supp}{\mathrm{supp}}
\newcommand{\indopp}{\bar{\mathbbm{1}}}
\newcommand{\ones}{\mathbbm{1}}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}

@@

\begin{theorem}[{\cite[Theorem 3.1.2]{weibel2007minkowski}}] \label{thm:unique-face-decomp}
	Let $T_1, \ldots, T_n$ be polytopes in $\reals^d$ and let $F$ be a face of the Minkowski sum $T := T_1 \oplus \ldots \oplus T_n$.
	Then there are faces $F_1, \ldots, F_n$ of $T_1, \ldots, T_n$ respectively such that $F = F_1 \oplus \ldots \oplus F_n$.
	Moreover, this decomposition is unique.
\end{theorem}

\iffalse 
\begin{corollary}[{\cite[Corollary 3.1.3]{weibel2007minkowski}}]\label{cor:face-decomp-normal-cones}
	Let $T = T_1 \oplus \ldots \oplus T_n$ be a Minkowski sum of polytopes in $\reals^d$, let $F$ be a nonempty face of $T$ , and let $F_1, \ldots, F_n$ be its decomposition.
	Then $\N(F;T) = \N(F_1;T_1) \cap \ldots \cap \N(F_n; T_n)$.
\end{corollary}

%used in proving completeness of constructed V
\begin{corollary}[{\cite[Corollary 3.1.4]{weibel2007minkowski}}]
	Let $F_1, \ldots, F_n$ be nonempty faces of the polytopes $T_1, \ldots, T_n$ respectively, then $F_1 \oplus \ldots \oplus F_n$ is a face of $T_1 \oplus \ldots \oplus T_n$ if and only if the intersection of their normal cones is nonempty.
\end{corollary}
\fi

\begin{theorem}[{\cite[Theorem 3.1.6]{weibel2007minkowski}}]\label{thm:support-minksum}
	The supporting function of a Minkowski sum is the sum of the supporting functions of its summands.
\end{theorem}

\cite{weibel2007minkowski} notes that:
\begin{quote}
	It is easy to see that the normal fan (undefined here, but consequently normal cones) of $p_i T_i$ does not change as long as $p_i$ is positive.  Since the normal fan of a Minkowski sum can be deduced from that of its summands, we can deduce from this that the conbimatorial properties of $\oplus_p T_y$ stay the same as long as all $p_i$ are positive.
\end{quote}

Suppose we are given a polytope $T_y \in \reals^d$ and set of vectors $V \in \reals^{k \times d}$.
Call $e^y \in \reals^k$ the vector such that $e^y_i = \max_{x \in T_y}\inprod{v_i}{x}$.  
For a finite set $\T = \{T_1, , \ldots, T_n\}$, let us denote the \emph{support matrix} $E = (e^y)_{y=1}^n$.
\begin{definition}\label{def:complete}
	We say a set of normals $V$ is \emph{complete} with respect to a polytope $T_y$ if $T_y = \{x \in \reals^d: Vx \leq e^y\}$.
\end{definition}
Moreover, we say $V$ is complete with respect to the set of polytopes $\T$ if and only if $V$ is complete with respect to each $T_y \in \T$.

@@

Now, for a given polytope $T(p)$, we want to ask when a given $z \in \reals^d$ is in the polytope $T(p)$.
We will later generalize to finding the set of $p \in \simplex$ for which $\vec 0 \in T(p)$ by substituting $z= \vec 0$.
Throughout, assume we have $V$ which is complete for $\T$ and $E$ defined by the support of each normal in $V$ for all $T_y \in \T$.
We denote $e^y = E_{;y}$ as the $y^{th}$ column of $E$, or equivalently, the support vector for $T_y$ given $V$.

@@

We want to know the following: starting from $\T$, can we derive the cell $C \subseteq \simplex$ where $\vec 0 \in T(p)$ for all $p \in C$?
We start with the latter, leaving the former for future work.

We know that if we are given $\T$ and a complete set of normals $V$, we can describe $D = \{p \in \simplex : Ep \geq \vec 0\}$.
However, we do not necessarily have $D' := \{p \in A : Ep \geq \vec 0\} \subseteq \simplex$, and this poses some issues given the construction of the cell $C$ only assuming that we are in the affine span of the simplex, but not necessarily in the simplex.

@@

\section{$d$-dimensional omitted proofs}\label{app:omitted-d-dim}
\begin{proof}[Proof of Lemma~\ref{lemma:minkowski-support}]
	\jessie{Appendix}
	For each $p$, $\oplus_p \T$ is a polytope.
	\bo{TODO: it turns out that $p$ doesn't matter, only its support; and there are only finitely many possible supports (or even better, the normals that are complete for the uniform minkowski sum turn out to be complete for all other $p$-combinations).}
	For each of the finitely many supports ($2^n - 1$), we know $\oplus_p \T$ is a polytope, and every polytope can be defined by a finite, complete set of vectors for that polytope.
	As a two polytopes with the same support are combinatorially equivalent\jessiet{is this what we want?  I know the idea but not sure if this is the proper term}, they can be defined by the same facet enumeration, and any set of normals that is complete for $\oplus_p \T$ is also complete for a $\oplus_{p'}\T$ if $\supp(p) = \supp(p')$.
	We can simply concatenate these finite set of normals for the finite polytope supports, with some normals possibly becoming redundant.
	This yields finitely many normals defining the weighted Minkowski sum $\oplus_p \T$ for all $p \in \simplex$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lemma:D-polytope}]
	\jessie{Appendix}
	Recall by definition, the notation $\oplus_p \T = \{\sum_y p_y x_y : x_y \in T_y (\forall y)\}$.
	Each $T_y$ is a polytope, so $p_y T_y$ is a polytope.
	The Minkowski sum of polytopes is a polytope, so $\oplus_p \T$ is a polytope by~\cite[Section 1.2]{weibel2007minkowski}.
	\bo{TODO}\jessie{Taking a try}
	Since $\oplus_p \T$ is a polytope for all $p \in \simplex$, we know there is a halfspace representation of normals $V$ so that for all $y \in \Y$, we have $x \in p_yT_y \iff \inprod{V}{x} \leq p_y e^y$ for some matrix $V$ and the support vector $e^y$, where $e^y_i = H_{T_y}(V_i)$.
	By Lemma~\ref{lemma:minkowski-support}, we know that there is a set of normals $V^*$ that is complete for $T(p)$ for all $p \in \simplex$.
	We construct $E^*$ as the support matrix for this complete set of normals.
    We know that the support of the Minkowski sum for a given normal is the sum of the normals~\cite[Theorem 3.1.6]{weibel2007minkowski}, and so we we can take the $x \in \oplus_p \T \iff \inprod{V^*}{x} \leq E^*p$.
	Substituting $x = \vec 0$, we see $\vec 0 \in \oplus_p \T \iff Ep \geq \vec 0 \iff p \in D(\T)$ by Lemma~\ref{lem:describe-D}, which defines a polytope by construction of $E$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lemma:E-to-B}]
%statement:
%	  Let $C = \{p : Bp \geq \vec 0 \}$ where $B$ has the minimum possible number of rows to capture $C$, and suppose $C = \{p : Ep \geq \vec 0 \}$.
%	Then for each row in $B$ there is some (unique) row in $E$ that is equal to $\alpha B$ for some positive $\alpha$.
	\bo{? Cite} \jessie{Is left as an exercise in the following text:} \cite[Exercise 2.15]{ziegler2012lectures} alludes to this fact.
	Suppose there was a row $j$ of $B$ that did not appear (possibly scaled, because of the inequality on $\vec 0$) in $E$.
	Then there is some $x \in \{x : Ex \geq 0\} = C$ so that $\inprod{B_i}{x} \geq 0$ for all $i \neq j$ and $\inprod{B_j}{x} < 0$, which contradicts $x \in \{x : x : Bx \geq 0\}$.
\end{proof}
